An Automatic Verification Framework for Software Configurations

### 2016-05-03 OSDI/abs.tex
config: abs and intro

This paper presents, \app, the first automatic verification framework for
general software configurations.
Given a target configuration file,
\app first analyzes a dataset containing sample configuration files,
thus generating well-structured and probabilistically-typed 
intermediate representations.
Then, \app derives rules and constraints by analyzing
the intermediate representations, thus building a language model.
Finally, \app uses the resulting language model
to verify the given configuration file.
\app framework is highly modular, 
does not rely on the system source code or templates and
can be applied to any new configuration file type with minimal user input.

\app is capable of detecting various tricky errors,
previous methods cannot detect,
including ordering errors, fine-grained value correlation errors, 
entry missing errors, and environment related errors. 
%Our tool, named \app, relies on an abstract representation of language 
%rules to allow for this modularity. 
\ennan{One or two sentences here describe the most important
experimental results. For example, we verify real-world misconfiguration 
files and find 95\% errors are detected.}

>>>>

This paper presents, \app, the first automatic verification framework for
general software configurations.
\app verifies a target configuration file $F$ through three steps.
First, \app analyzes a dataset containing many sample configuration 
files belonging to the same system as $F$,
and translates these sample files to a
well-structured and probabilistically-typed 
intermediate representation.
Second, \app derives rules and constraints by analyzing
this intermediate representation, thus building a
language model.
Finally, \app uses the resulting language model to verify $F$.
\app framework is highly modular, 
does not rely on the system source code or templates and
can be applied to any new configuration file type with minimal user input.

\app is capable of detecting various tricky errors that cannot
be detected by previous efforts,
including ordering errors, fine-grained value correlation errors, 
entry missing errors, and environment related errors. 
%Our tool, named \app, relies on an abstract representation of language 
%rules to allow for this modularity. 
\ennan{One or two sentences here describe the most important
experimental results. For example, we verify real-world misconfiguration 
files and find 95\% errors are detected.}

### 2016-05-03 OSDI/intro.tex
config: abs and intro

Configuration errors are one of the most important root causes of
today's software system failures~\cite{xu15systems, yin11anempirical}.
In practice, misconfiguration problems may lead to security
vulnerabilities, application crashes, severe disruptions in software
functionality, and incorrect program executions%
~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.  
For example, Yin {\em et al.}~\cite{yin11anempirical} reveal
that about 31\% system failures are caused by misconfiguration problem.

Although many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{wang04automatic, attariyan10automating,
su07autobash, whitaker04configuration}, 
offering automatic verification to configuration files -- like what we
did for verifying regular programs -- is still an open problem.
The main reasons for this are:
1) entries in configuration files are untyped assignments, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

@@

This paper presents \app, the first automatic verification framework
for general software configurations.
The insight of \app design is that we first analyze and learn 
a dataset sample of configuration files, thus building 
a structured and typed language model. Using this ``learned'' language
model, we can verify the target configuration files.
Such a design insight overcomes the above obstacles:
1) learned language model itself is well-structured and typed, so
that we can parse target configuration files to it and verify these
files; 2) because we learn and derive rules and constraints 
from an existing sample dataset, 
we do not need users to write explicit constraints or specifications.

In general, \app's workflow to verifying a given configuration file
could be looked as a three-step methodology.
First, \app analyzes a dataset containing sample configuration files,
thus generating a well-structured and probabilistically-typed 
intermediate representation.
Second, \app derives rules and constraints by analyzing
the intermediate representations, thus building a language model.
Finally, \app uses the resulting language model
to verify the given configuration file and detect potential errors.

Building such an automatic verification framework for
configuration files, nevertheless, still requires addressing several
challenges. 
First, when inferencing each entry's type in a given configuration file, 
the type of a variable cannot always be fully determined from a single 
value. For example, an entry {\tt foo = MAX\_SIZE} is most likely
an integer rather than string. We address this problem by introducing 
so called {\emph{probabilistic types}}.
Rather than assigning only one variable to a single type,
like previous work did~\cite{zhang14encore}, 
we assign several types with their probability distributions. 
The entry in the above example should be assigned 
a probabilistic type like 
{\tt \{foo, MAX\_SIZE, [(Int, 90\%), (String, 10\%)]\}}.
\ennan{I am not quite sure probabilistic types are obviously
better than deterministic types.}

Second, \ennan{Here, we need to add a challenge description about
why using learning algorithm is interesting.}

@@

We can then use these more structured files
as a training set to learn the rules. The learning algorithm
is template-based to be easily extensible. We provide an initial set of templates and the
learner learns some concrete instances from the training set. These
rules are used for detecting errors violating the learned constraints
in the files given by the user.

As an 
illustration of a simple rule that we can learn, consider a template
 $X_1 \le X_2$, where $X_1$ and $X_2$ are
integer variables. The learner might derive the rule stating that
$\texttt{mysql.max\_persistent} \le \texttt{max\_connections}$. There is a classification and taxonomy of configuration errors in the 
existing work on automated configuration troubleshooting~\cite{yin11anempirical, configdataset}. We provide templates for every class that \app can handle: we consider integer constraints, ordering
constraints, typing constraints, and constraints about correlated entries (such as ``if $X$ is present, $Y$ has to appear as well''). 
Unfortunately, we cannot handle the class of errors that rely on the analysis of the whole operating system.
Our language-based approach can only learn on sets of text files, not the system environment.

From a practical perspective \app introduces no additional burden 
to the users: they can simply use \app to check for errors in their configuration files. However, they can also easily extend the framework themselves. The system is designed to be highly modular. If there is a class of rules that \app is not currently learning, the user can develop their own templates and learners for that class. The new learner can be added to \app and this way it can check additionally a new set of errors.

\com{
Finally, from a systems perspective this is the first approach that {\emph{proactively}} checks 
 the correctness of configuration files. All previous work
~\cite{xu15systems,zhang14encore,yuan11context, wang04automatic,attariyan10automating,
su07autobash,whitaker04configuration} tries to identify the problem after the
failure occured. Our approach isolates potential errors before the system failure occurs, e.g. before the installation. We can also see \app as a tool that can run in conjunction with existing tools. Pre-analyzed configuration files are already free from language-based errors, and this way the workloads of post-failure forensics at the runtime
is significantly reduced, thus making these tools truly practical.
}

@@

  \item We designed and implemented a tool, \app, that can learn a
language model from an example set of correct configuration files, and
we use the model to verify new configuration files.
  \item We use probabilistic types to assign a confidence distribution over a set of types to a value.
  \item In \app we define a interface for describing a verification attribute in a learning context, making it easy to add new rules to the system.

>>>>

Configuration errors are one of the most important root causes of
today's software system failures~\cite{xu15systems, yin11anempirical}.
For example, Yin {\em et al.}~\cite{yin11anempirical} reveal
that about 31\% system failures were caused by misconfiguration problems. 
Misconfigurations, in practice, may result in various
system-wide problems, such as security vulnerabilities, 
application crashes, severe disruptions in software
functionality, and incorrect program executions%
~\cite{zhang14encore, yuan11context, xu13do, xu15hey}.

While many efforts have been proposed 
to check, troubleshoot, diagnose and repair configuration errors%
~\cite{attariyan10automating,
su07autobash, whitaker04configuration}, 
offering automatic verification to configuration files -- like what we
did for verifying regular programs -- is still highly
desirable~\cite{wang04automatic, zhang14encore}.
Nevertheless, the main obstacles are:
1) entries in configuration files are untyped entries, 2) there
is no explicit structure policy for the entries in configuration
files, and 3) there are surprisingly few rules specifying the
entries' constraints.

In order to avoid these obstacles,
researchers propose statistical analysis and learning based approaches%
~\cite{wang04automatic, zhang14encore, yuan11context}. 
These efforts build checking policies by learning a sample 
data set, rather than explicitly specifying entries' types or rules.
In particular, for each entry in a certain configuration file, 
if it deviates from the common values used in a large collection
of configurations (\ie, the training set), it is typically
suspected as a potential configuration error.
However, because these learning 
efforts {\em either} are limited to simplistic 
configuration errors (\eg, type errors and syntax errors), 
{\em or} heavily rely on template-based inference~\cite{zhang14encore}, 
many sophisticated configuration errors, 
\eg, hard to be templated in practice, cannot be detected.
For example, if {\tt extension = mysql.so} appears 
before {\tt extension = recode.so} in PHP configuration file,  
it would lead to a crash error, since the correct ordering 
should be {\tt extension = recode.so} before 
{\tt extension = mysql.so}~\cite{yin11anempirical};
however, such an error cannot be detected by existing
learning efforts, since it is hard to build a template for this
error~\cite{xu15systems}.

@@

In order to truly achieve automatic configuration verification,
we argue that we need to leverage a collection of more powerful learning 
algorithms (not necessarily depending on templates) to derive 
a more sophisticated language model (with abundant rules covering
tricky configuration errors), and then check the configuration files
of interest through the learned language model.

Based on the above insight, 
this paper presents \app, the first automatic verification framework
for general software configurations.
In particular, \app's workflow to verifying a given configuration file
could be looked as a three-step methodology.
First, \app analyzes a dataset containing sample configuration files,
thus generating a well-structured and probabilistically-typed 
intermediate representation.
Second, \app derives rules and constraints by analyzing
the intermediate representations, thus building a language model.
Finally, \app uses the resulting language model
to verify the given configuration file and detect potential errors.
Compared with previous efforts,
\app does not necessarily rely on users' templates, 
and is capable of detecting more tricky configuration errors that
cannot be identified by existing work (details in $\S$\ref{sec-motiv}).

Building such an automatic verification framework for
configuration files, nevertheless, requires addressing several challenges. 
First, before formulating the language model, 
we need to infer each entry's type in a given configuration file;
however, the type of a variable cannot always be fully determined 
from a single value. 
For example, an entry {\tt foo = MAX\_SIZE} is most likely
an integer rather than string. We address this problem by introducing 
so called {\emph{probabilistic types}}.
Rather than assigning only one variable to a single type,
like previous work did~\cite{zhang14encore}, 
we assign several types with their probability distributions. 
The entry in the above example should be assigned 
a probabilistic type like 
{\tt \{foo, MAX\_SIZE, [(Int, 90\%), (String, 10\%)]\}}.
\ennan{I am not quite sure probabilistic types are obviously
better than deterministic types.}

Second, without template, how to learn rules and constraints present
a difficult. \ennan{Here, we need to add a describe to illustrate
how we learn rules without templates.}

@@

From a practical perspective, 
\app introduces no additional burden 
to the users: they can simply use \app to check for errors in their
configuration files. However, they can also easily extend the framework
themselves. The system is designed to be highly modular. If there is a
class of rules that \app is not currently learning, the user can develop
their own templates and learners for that class. The new learner can be
added to \app and this way it can check additionally a new set of
errors.

Our \app prototype still has a few limitations,
\eg, cannot handle errors occurred in execution time.
Nevertheless, we believe \app may suggest a practical path
toward automatic and language-based configuration verification.

\com{
Finally, from a systems perspective this is the first approach that {\emph{proactively}} checks 
 the correctness of configuration files. All previous work
~\cite{xu15systems,zhang14encore,yuan11context, wang04automatic,attariyan10automating,
su07autobash,whitaker04configuration} tries to identify the problem after the
failure occurred. Our approach isolates potential errors before the system failure occurs, e.g. before the installation. We can also see \app as a tool that can run in conjunction with existing tools. Pre-analyzed configuration files are already free from language-based errors, and this way the workloads of post-failure forensics at the runtime
is significantly reduced, thus making these tools truly practical.
}

@@

\item We propose the first automatic configuration verification
framework, \app, that can learn a language model from an example set of 
correct configuration files, and we use the model to verify 
new configuration files.

\item \app uses probabilistic types to assign a confidence 
distribution over a set of types to each entry, 
while generating the intermediate representation.

\item \app applies existing machine learning algorithms to 
enable powerful rule and constraint inference without the assistance 
from any pre-defined templates.

\item \app is capable of detecting various tricky errors that cannot
be detected by previous efforts,
including ordering errors, fine-grained value correlation errors, 
entry missing errors, and environment related errors.

\item We implement a \app prototype and evaluate it by
conducting comprehensive experiments.
